\documentclass[12pt,twoside,a5paper]{article}

\usepackage{multicol}

\usepackage[main=english]{babel}
\usepackage{divine-office}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\begin{document}

\title{Lelya --- Sundays}
\author{}
\date{}
\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\begin{halfparskip}
  \liturgicalhint{For singing:}

  \cc~Glory be given / unto our God / in the heights of heaven.\\
  Peace be on earth / and hope to mankind / in the heights of heaven.

  \rr~Amen, bless, my lord. \liturgicalhint{Kiss of peace.}

  Heav'nly Father of us all / may Your Name be ever blest / \\
  may Your kingdom come to all / Holy, holy are You, Lord!

  Heav'nly Father of us all / Your glory fills heav'n and earth / \\
  Men and angels cry aloud / Holy, holy are You, Lord!

  Heav'nly Father of us all / may Your Name be ever blest /\\
  may Your will be done on earth / as it always is in heav'n.

  Grant today the bread we need / pardon us our debts and sins /\\
  as we have forgiven those / who somehow offended us.

  Do not let us fall away / when temptation comes to sin / \\
  but deliver us always / save us from the evil one!

  For the kingdom, Lord, is Yours / and the power and glory / \\
  for ever and evermore / Amen; evermore; amen.

  \cc~Glory to the Father be / Son and Spirit, equally / \\
  from eternity and now / and forevermore. Amen.

  Heav'nly Father of us all / Your glory fills heav'n and earth / \\
  Men and angels cry aloud / Holy, holy are You, Lord!

  Heav'nly Father of us all / may Your Name be ever blest / \\
  may Your will be done on earth / as it always is in heav'n.
\end{halfparskip}

\begin{halfparskip}
  \liturgicalhint{For recitation:}

  \cc~Glory to God in the highest (3x). And on earth peace and good hope to men at all times for ever. [Amen]\footnote{The Hudra and the Syro-Malabar liturgy add here ``Amen'', contrary to Breviarium.}.

  \rr~Bless, my Lord. \liturgicalhint{Kiss of peace.}

  Our Father in heaven, hallowed be Your Name; Your kingdom come; holy, holy, holy are You; our Father in heaven, heaven and earth are full of the greatness of Your glory; angels and men cry out to you: Holy, holy, holy are You.

  Our Father in heaven, hallowed be Your Name; Your kingdom come, Your will be done on earth as it is in heaven. Give us this day the bread we need; and forgive us our offences and sins as we have forgiven those who have offended us. Lead us not into temptation, but deliver us from the evil one. For Yours is the kingdom, the power and the glory for ever and ever. Amen.

  \cc~Glory be to the Father and to the Son and to the Holy Spirit.

  From eternity and forever. Amen and amen. Our Father in heaven, hallowed be Your Name; Your kingdom come. Holy, holy, holy are You. Our Father in heaven, heaven and earth are full of the greatness of Your glory; angels and men cry out to You: Holy, holy, holy are You.
\end{halfparskip}

\dd~Let us arise to prayer, let us pray, peace be with us.

\cc~May we arise, my Lord, in the hidden power of Your Godhead, let us be confirmed (\translationoption{strengthened}) in the wondrous hope of Your Majesty, let us be lifted up and strengthened by the high arm of Your might; and may we be worthy by the help of Your grace to lift up to You praise, honour, confession and worship, at all times, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Psalms}

\liturgicalhint{On Sundays 3 hulale are recited farced with alleluia (respectively hulale 5-7 ; or 9-11 + Ps 81 ; or 12-14 or 16-18 + Ps 129)\footnote{In this Lelya booklet we replace every hulala by a marmita, excluding the marmiyata used in ramsa. We follow the ancient system in which the last hulala (here marmita) is ended by the psalm accompanying the qalta: pss 81 \& 129.}, followed on Sundays 'before' by Ps. 81, on Sundays 'after' by Ps. 129, farced with alleluia after each clause.}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsunday{1. Sundays ``before''}{Marmita 29 (Pss 75-77)}

\begin{halfparskip}
  \Psalm{75} We praise You, O God;~\sep\ we praise You and call on Your Name.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). -- First verse.}

  We have related all Your wonderful deeds.~\sep\ ``For I will receive time, and I will judge with uprightness''.~\sep

  The earth will be humbled, and all its inhabitants;~\sep\ You have established its inhabitants.~\sep

  You said to the foolish, ``Deal not foolishly'',~\sep\ and to the wicked, ``Do not lift up the horn;~\sep

  do not lift up on high Your horn~\sep\ or speak haughtily'' (\translationliteral{with a haughty neck}).~\sep

  For there is no going forth from the west~\sep\ or from the desert of the mountains.~\sep

  For God is the judge;~\sep\ He humbles one and He exalts another.~\sep

  For a cup is in the hand of the Lord;~\sep\ it is full of a mixture of wine-dregs.~\sep

  He lowered (\translationoption{inclined}) it from one to another;~\sep\ all the wicked of the earth will strain out its dregs and drink them.~\sep

  But I will live forever;~\sep\ and I will sing to the God of Jacob.~\sep

  I will hack into pieces all the horns of the wicked;~\sep\ but the horns of the righteous will be lifted up.
\end{halfparskip}

\begin{halfparskip}
  \Psalm{76} God is known in Judah;~\sep\ His Name is great in Israel.~\sep

  His tabernacle will be in Salem;~\sep\ His dwelling place in Zion.~\sep

  There He shattered the arms of bows,~\sep\ the shield and the sword in battle.~\sep

  You are more resplendent and glorious than Your strong (\translationoption{fortified}) mountain.~\sep

  All the foolish-hearted were troubled;~\sep\ the mighty men have slept their sleep.~\sep

  Their hands were incapable because of Your rebuke, O God of Jacob;~\sep

  The riders of horses fell into a deep sleep, and~\sep\ You are fearsome.~\sep

  Who can stand before You in this anger?~\sep\ You heard judgment from the heavens;~\sep

  the earth saw and feared,~\sep\ when God was arising in order to judge~\sep\ and to save all the poor of the earth.~\sep

  For the mind of man will acknowledge You;~\sep\ and the remainder of His wrath will turn away anger.~\sep

  Vow and repay to the Lord your God;~\sep\ may all those around Him bring gifts to the awesome One,~\sep

  who humbles the spirit of rulers (\translationoption{princes})~\sep\ and is feared by the kings of the earth.
\end{halfparskip}

\begin{halfparskip}
  \Psalm{77} I cried out to God with my voice, and He heard me;~\sep\ I lifted up my voice to Him, and He answered me.~\sep

  In the day of my distress I sought the Lord.~\sep\ His hand beat me during the night, and I could not be still;~\sep\ there was no one to comfort my soul.~\sep

  I remembered God, and I was disturbed;~\sep\ and I thought, and my spirit was weary (\translationoption{agitated}).~\sep

  Dizziness took hold of my eyes;~\sep\ I became mute and did not speak.~\sep

  I thought about the days of old;~\sep\ I remembered the years of long ago.~\sep

  I meditated during the night and reflected in my heart (\translationoption{mind})~\sep\ I searched my spirit and said,~\sep

  Has the Lord forgotten me forever?~\sep\ Will He no more be pleased with me?~\sep

  Or will He forever remove His grace?~\sep\ Will He fulfil His word throughout the generations?~\sep\ Or has God forgotten to be compassionate?~\sep

  Or is He withholding His mercies in His anger?~\sep

  I said, ``It is my infirmity''~\sep\ They are a repetition [of strokes] of the right hand of the Most High.~\sep

  For I have remembered Your wondrous deeds of long ago.~\sep

  I have meditated on all Your works;~\sep\ I have pondered on Your artifices (\translationoption{doings}).~\sep

  O God, Your way is holy;~\sep\ there is no one great like our God.~\sep

  You are the God who has accomplished wonders~\sep\ and You have displayed Your strength among the peoples.~\sep

  You have redeemed Your people by Your arm,~\sep\ the sons (\translationoption{descendants}) of Jacob and Joseph.~\sep

  The waters saw You, O God; the waters saw You and feared.~\sep\ The depths also trembled, and the clouds dropped water;~\sep

  the heavens of the heaven cried out (\translationliteral{gave a voice}).~\sep\ Your arrows also were flying.~\sep

  The sound (\translationoption{voice}) of Your thunders was in the circuits (\translationoption{spheres});~\sep\ Your lightnings illumined the world.~\sep\ The earth was disturbed and trembled.~\sep

  Your path is in the sea; and Your ways are in the great waters.~\sep\ Your footprints were not perceived (\translationoption{known}).~\sep

  You led Your people like sheep (\translationoption{a flock})~\sep\ by the hand of Moses and Aaron.
\end{halfparskip}

\markedsunday{Sundays ``after''}{Marmita 49 (118b)}

\begin{halfparskip}
  \Psalm{118b} \acrosticletter{Lamadh} You, O Lord, are forever; and Your word stands firm in the heavens.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). -- First verse.}

  Your faithfulness is to generations;~\sep\ You established the earth and furnished it.~\sep

  But for Your law, which has been my meditation, I would perish in my humiliation.~\sep

  I will never forget Your commandments, for my life is in them.~\sep

  I belong to You; deliver me, for I have kept Your commandments,~\sep

  The wicked have waited for me so that they might destroy me, but I have understood Your testimony.~\sep

  I have seen an end to every uttermost part; but Your commandment is very broad.~\sep

  \acrosticletter{Mim} How I love Your law! All day it has been my meditation.~\sep

  Make me wiser than my enemies, for I have kept Your commandments.~\sep

  Give me more understanding than all my teachers, for Your testimony is my meditation.~\sep

  I have more understanding than the elders, for I have kept Your commandments.~\sep

  I have withheld my feet from every evil way,~\sep\ so that I might keep Your commandments.~\sep

  I have not departed from Your judgments, for You have taught me.~\sep

  Your words are sweeter to the roof of my palate than honey to my mouth.~\sep

  I have meditated on Your commandments; therefore I hate the entire path of the wicked.~\sep

  \acrosticletter{Nun} Your word is a lamp to my feet and a light to my paths.~\sep

  I have taken an oath and confirmed it, that I will keep Your righteous judgments.~\sep

  I have been greatly humbled; O Lord, restore my life according to Your word.~\sep

  Be pleased with the words of my mouth, O Lord; and teach me from Your judgments.~\sep

  My soul is always in Your hands; and I have not forgotten Your law.~\sep

  Sinners have placed traps for me; yet I have not departed from Your commandments.~\sep

  I have inherited Your testimony forever, for it is the delight of my heart.~\sep

  I have turned my heart, so that I might practice Your commandments forever in truth.~\sep

  \acrosticletter{Simkath} I hate the wicked, but I love Your law.~\sep

  You are my hiding place and my refuge; and I have hoped in Your word.~\sep

  Turn aside from me, wicked ones, so that I may keep the commandments of my God.~\sep

  Confirm me in Your word and I will live; and do not put me to shame because of my hope.~\sep

  Uphold me, and I will be saved; and I will be instructed in Your commandments.~\sep

  I have always rejected (\translationoption{abhorred}) all those who turn aside (\translationoption{err}) from You,~\sep\ for their meditation is wicked.~\sep

  Sustain me and I will be saved; and I will always meditate on Your commandments.~\sep

  My flesh creeps from fear of You; and I stand in awe of Your judgments.~\sep

  \acrosticletter{`E} Doer of justice (\translationoption{judgment}) and righteousness, leave me not in the hand of my oppressors.~\sep

  Delight Your servant with good things, so that the proud may not prevail over (\translationoption{oppress}) me.~\sep

  My eyes look for Your salvation, and for Your righteous promise.~\sep

  Deal with Your servant according to Your mercies; and teach me Your law.~\sep

  I am Your servant; give me understanding, so that I may understand (\translationoption{know}) Your testimony.~\sep

  It is time to serve the Lord. But look! They have made void Your law.~\sep

  Therefore I love Your commandments more than gold, and more than precious stones.~\sep

  I love all Your commandments; and I hate the entire path of the wicked.~\sep

  \acrosticletter{Pe} Your testimonies are great; therefore I keep them.~\sep

  Open Your word and give light; and grant understanding to the little children.~\sep

  I opened my mouth and drew breath; and I looked for Your salvation.~\sep

  Turn to me and have mercy on me, for I love Your Name.~\sep

  Establish my steps in Your paths; and may the wicked not prevail against me.~\sep

  Save me from the oppression of man, so that I may keep Your commandments.~\sep

  Cause Your face to shine on Your servant; and teach me Your law.~\sep

  Streams of water come from my eyes, because they have not kept Your law.~\sep

  \acrosticletter{Sadhe} You are righteous, O Lord; and very upright are Your judgments.~\sep

  You commanded Your testimony in righteousness and faithfulness (\translationoption{truth}).~\sep

  Zeal has always tortured me, for Your enemies have forgotten Your word.~\sep

  Your word is very pure; and Your servant loves it.~\sep

  I am small and despised; yet I have not forgotten Your commandments.~\sep

  Your righteousness remains forever; and Your law is in truth.~\sep

  Affliction and grief have come on me; but I have meditated on Your commandments.~\sep

  Your testimony is righteous forever; give me understanding, and I will live.~\sep

  \acrosticletter{Qoph} I cried to You with all my heart, Lord, hear (\translationoption{answer}) me, and I will keep Your precepts.~\sep

  I have called out to You; save me, and I will keep Your testimony.~\sep

  I rose early in the morning and called out; and looked (\translationoption{waited}) for Your word.~\sep

  My eyes preceded the (night)watch, so that I might meditate on Your word.~\sep

  Hear my voice, O Lord, according to Your mercies; and restore me to life by Your judgments.~\sep

  Those who wickedly persecute me draw near; and they are far removed from Your law.~\sep

  You are near, O Lord; and all Your commandments are in truth;~\sep

  my eyes preceded (\translationoption{prevented}) the (night)watch, so that I might meditate on Your word.~\sep

  Of old have I know Your testimony, for You established it from everlasting.~\sep

  \acrosticletter{Rish} Look on my humiliation and deliver me, for I have not forgotten (\translationoption{disregarded}) Your law.~\sep

  Judge my cause (\translationoption{judgment}) and deliver me; restore my life by Your word.~\sep

  Salvation is far from the wicked, for they have not sought Your commandments.~\sep

  Your mercies are many, O Lord; and restore my life by Your judgments.~\sep

  My persecutors and my enemies are numerous; but I have not declined from Your testimony.~\sep

  I have seen the wicked; and I know that they have not kept Your word.~\sep

  See that I love Your commandments, O Lord; restore my life in Your lovingkindness.~\sep

  The beginning of Your word is truth; and all Your righteous judgments are forever.~\sep

  \acrosticletter{Shin} Nobles have persecuted me without cause; and my heart stands in awe of Your word.~\sep

  I rejoice in Your word, like one who finds great spoil.~\sep

  I hate and abhor iniquity; but I love Your law.~\sep

  Seven times a day I praise You, because of Your righteous judgments.~\sep

  Great is the peace of those who love Your law; there is no infirmity in them.~\sep

  I have put my hope in Your salvation, O Lord; and I have practised Your commandments.~\sep

  My soul has kept Your testimony;~\sep\ and loved it exceedingly.~\sep

  I have kept Your commandments and Your testimonies; and all my ways are before You.~\sep

  \acrosticletter{Taw} May my song enter before You, O Lord;~\sep\ quicken me by Your word.~\sep

  May my petition enter before You; and deliver me by Your word.~\sep

  May my tongue pour forth Your word [a hymn],~\sep\ for all Your commandments are in righteousness.~\sep

  May my lips speak Your praises, while You teach me Your commandments.~\sep

  May Your hand help me, for I delight in Your commandments.~\sep

  My soul has looked for Your salvation; and I have meditated in Your law.~\sep

  May I live and praise You; as for Your judgment, it will help me.~\sep

  I have wandered like a lost sheep. Seek Your servant,~\sep\ for I have not forgotten Your precepts.

  Glory... --- Alleluia, alleluia, glory to You, O God, alleluia, alleluia, glory to You, O God, alleluia, alleluia, O Lord, have mercy on us.
\end{halfparskip}

\dd~Let us pray; peace be with us.

\cc~Strengthen, our Lord, and our God, our weakness in Your compassion; encourage (\translationoption{comfort}) and help the feebleness of our soul in Your grace; wake up the drowsiness of our minds; lighten (\translationoption{take away}) the burden of our limbs; wash and cleanse the filth of our debts and sins; illumine the darkness of our intellects; stretch out and give us strength and a helping hand, that thereby we may rise to confess and glorify You unceasingly all the days of our life, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsunday{2. Sundays ``before''}{Marmita 30 (Ps 78)}

\begin{halfparskip}
  \Psalm{78} My people, give ear (\translationoption{give heed}) to my law and be obedient,~\sep\ and incline Your ears to the words of my mouth.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). --- First verse.}

  For look! I will open my mouth in parables;~\sep\ I will utter allegories of a long time ago.~\sep

  We know those things that we have heard,~\sep\ and also those things that our ancestors (\translationoption{fathers}) related to us~\sep

  so that we might not hide (them) from their children,~\sep\ but relate to another generation~\sep

  the praises of the Lord and His power~\sep\ and His wondrous deeds that He has accomplished.~\sep

  He has established His testimony in Jacob;~\sep\ and He has laid His law in Israel,~\sep

  as He commanded our ancestors~\sep\ that they should make (it) known to their children,~\sep

  so that another generation might know (it):~\sep\ the children who are being born and are remaining (\translationoption{arise}),~\sep\ so that they might relate (it) to their children,~\sep

  so that their hope might be in God~\sep\ and that they not forget the works of God,~\sep\ but keep His commandments.~\sep

  And that they not be like their ancestors (\translationoption{fathers}),~\sep\ an evil and contentious (\translationoption{provoking}) generation,~\sep\ a generation that did not set its heart in order~\sep\ and did not believe the God of their spirit.~\sep

  The sons of Ephraim who stretched the bowstring and shot with the bow,~\sep\ but turned back in the day of battle,~\sep

  because they did not keep the covenant of God~\sep\ and did not wish to walk in His law.~\sep

  They forgot His works and His wondrous deeds~\sep\ that He showed them in the presence of their fathers (\translationoption{ancestors}).~\sep

  For He did wondrous deeds in the land of Egypt~\sep\ and in the fields of Zoan.~\sep

  He made a breach in the sea and made them cross over;~\sep\ and He caused the waters to stand as in bottles.~\sep

  He led them during the day by a cloud,~\sep\ and all night by a light of fire.~\sep

  He cleft the rock in the wilderness;~\sep\ and He gave them drink as from the great deep.~\sep

  He brought forth streams (\translationoption{running water}) from the rock;~\sep\ the waters flowed like rivers.~\sep

  But the people continued to sin (\translationoption{sinned still more});~\sep\ and to provoke the Most High in the thirsty region.~\sep

  They tested God in their heart~\sep\ by asking for food for their souls.~\sep

  They murmured against God and said,~\sep\ ``Is God able to spread tables for us in the wilderness?~\sep

  If He struck the rock and waters flowed and torrents shed forth,~\sep\ is He able to give us bread also? Or can He provide food for His people?''~\sep

  Because of this the Lord heard and became angry;~\sep\ and a fire was kindled against Jacob;~\sep\ anger also rose against Israel.~\sep

  Because they did not believe in God;~\sep\ they did not hope for His salvation.~\sep

  (B) He commanded the clouds from above;~\sep\ and the doors of the heavens were opened.~\sep

  He made manna come down for them to eat;~\sep\ and He gave the bread of heaven to them.~\sep

  Men ate the bread of angels;~\sep\ and He sent them food to satisfy them.~\sep

  He made winds blow in the heavens;~\sep\ and by His strength He guided the south wind.~\sep

  He brought down to them flesh like dust~\sep\ and feathered birds like the sand of the seas.~\sep

  They fell in their encampments, all around their tents.~\sep

  They ate and were very satisfied; and He gave them what they craved.~\sep

  But they did not withdraw from their cravings;~\sep\ and while their food was (still) in their mouths,~\sep

  the anger of God rose against them.~\sep\ and He slew some of their wealthy;~\sep\ He made the chosen ones of Israel to kneel down.~\sep

  In all these things they sinned;~\sep\ and again they did not believe in His wondrous works.~\sep

  They ended their days in vanity,~\sep\ and their years in haste.~\sep

  When He slew them, they would seek Him;~\sep\ and they would turn back and come to Him.~\sep

  They remembered that God was their Helper~\sep\ and the Most High God was their Saviour.~\sep

  They loved Him with their mouth, but with their tongue they lied to Him.~\sep

  They did not make their heart right with Him; and they did not believe in His covenant.~\sep

  Yet He was merciful, forgiving sins and not destroying.~\sep\ Time and again He turned away His anger and did not arouse all His wrath.~\sep

  He remembered that they were flesh;~\sep\ a wind that passes (\translationoption{goes}) and does not return.~\sep

  They provoked Him in the wilderness;~\sep\ and they angered Him in the desert.~\sep

  They turned away and tested God;~\sep\ and provoked the Holy One of Israel.~\sep

  They did not remember His hand, the day that He saved them from the oppressor,~\sep

  that He performed His signs in Egypt~\sep\ and His wonders in the fields of Zoan.~\sep

  For He turned their rivers into blood,~\sep\ and their streams, so that they could not drink the water.~\sep

  He sent upon them a swarm (of insects) that consumed them,~\sep\ and frogs that devastated them.~\sep

  He gave their harvests to the locusts,~\sep\ and their labour to the crawling (\translationoption{young}) locust.~\sep

  He ruined their vines with hail~\sep\ and their fig trees with ice.~\sep

  He gave their cattle over to the hail~\sep\ and their possessions to the fire (\translationoption{conflagration}).~\sep

  He sent upon them His fierce anger (\translationoption{the fierceness of His anger}).~\sep\ Wrath and anger and distress~\sep\ He sent by the hand of an evil angel.~\sep

  He forced open evil paths;~\sep\ and He did not spare their souls from death.~\sep\ He delivered their animals over to death.~\sep

  He slew all the firstborn of Egypt;~\sep\ the chief of all their progeny in the tabernacle of Ham.~\sep

  (C) He led forth His people like sheep;~\sep\ He led them like a flock in the wilderness.~\sep

  He made them dwell in hope, and they did not fear;~\sep\ but the sea covered (\translationoption{overwhelmed}) their enemies.~\sep

  He brought them to the border of His sanctuary;~\sep\ to this mountain that His right hand acquired.~\sep

  He destroyed the peoples from before them;~\sep\ He placed them by the lot (\translationliteral{measuring lines}) of His inheritance.~\sep\ The tribes of Israel dwelt in their tents.~\sep

  They tested and provoked the Most High God;~\sep\ and did not keep His testimonies.~\sep

  They turned away and acted deceitfully like their fathers (\translationoption{ancestors});~\sep\ they were turned away like a deceitful (\translationoption{crooked}) bow.~\sep

  They angered Him with their high places; and provoked Him to jealousy with their idols.~\sep

  God heard and became furious;~\sep\ He was very angry with Israel.~\sep

  He forgot (\translationoption{abandoned}) the tabernacle of Shiloh,~\sep\ the tabernacle where He dwelt among men.~\sep

  He delivered (\translationoption{gave}) His people to captivity; and His glory into the hand of the oppressor.~\sep


  He delivered His people to the sword; and disregarded (\translationoption{turned away from}) His inheritance.~\sep

  Fire consumed their young men;~\sep\ and their young women were ill-treated.~\sep

  Their priests fell by the sword;~\sep\ and their widows were not lamented.~\sep

  The Lord awoke as from sleep,~\sep\ and like a mighty man who has become sober (\translationliteral{shaken off his wine}).~\sep

  He struck His enemies in the back (\translationoption{behind Him});~\sep\ He made (\translationliteral{give}) them the reproach of the world.~\sep

  He rejected the tabernacle of Joseph;~\sep\ and was not pleased with the tribe of Ephraim.~\sep

  He chose the tribe of Judah;~\sep\ mount Zion, which He loved.~\sep

  He built His sanctuary on a high place;~\sep\ and established it on the earth forever.~\sep

  He chose David His servant;~\sep\ and led him away from the sheepfold~\sep

  and from behind the nursing sheep,~\sep\ so that He might shepherd (\translationoption{fed}) Jacob His people and Israel His inheritance.~\sep

  So He shepherded (\translationoption{fed}) them in the integrity of His heart;~\sep\ by the skill (\translationoption{understanding}) of His hands He led them.
\end{halfparskip}

\markedsunday{Sundays ``after''}{Marmita 50 (119)}

\begin{halfparskip}
  \Psalm{119} I called out to the Lord in my affliction; and the Lord answered me and delivered my soul.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). --- First verse.}

  from the lips of the wicked~\sep\ and from deceitful tongues.~\sep

  What will they give you, and what will they add to you,~\sep\ the deceitful tongues?~\sep

  The arrows of a man are sharp,~\sep\ like the burning coals of an oak.~\sep

  Woe to me that my pilgrimage (\translationoption{sojourn, exile}) has lasted (\translationoption{is prolonged}); and I dwell in the tent of Kedar!~\sep

  My soul has dwelt for a long time with those who hate peace.~\sep

  I was speaking peace,~\sep\ but they were fighting with me.~\sep

  \Psalm{120} I will lift my eyes to the mountains.~\sep\ Where will my help come from?~\sep

  My help is from the Lord,~\sep\ who made the heavens and the earth.~\sep

  He will not allow (\translationliteral{give}) your foot to slip;~\sep\ the One who watches over you will not slumber.~\sep

  For He does not slumber, nor does He sleep,~\sep\ He who watches over Israel.~\sep

  The Lord is your watchman;~\sep\ the Lord will shade (\translationoption{cover}) you with His right hand.~\sep

  By day the sun will not harm you,~\sep\ nor the moon by night.~\sep

  The Lord will protect you from all evils;~\sep\ the Lord will watch over your soul.~\sep

  He will watch over your going out and your coming in,~\sep\ from now and forever.~\sep

  \Psalm{121} I was glad when they were saying to me,~\sep\ We are going to the house of the Lord.~\sep

  My feet were standing at your gates, Jerusalem.~\sep

  Jerusalem is built~\sep\ like a city that is surrounded by a wall.~\sep

  The tribes went up there, the tribes of the Lord,~\sep\ as a testimony to Israel, to give thanks to the Name of the Lord.~\sep

  For there they set up thrones for judgment,~\sep\ thrones of the house of David.~\sep

  Pray (\translationliteral{ask}) for the peace of Jerusalem;~\sep\ may those who love You be prosperous.~\sep

  May there be peace with your army (\translationoption{strength})~\sep\ and prosperity within your palaces (\translationoption{walls}).~\sep

  For the sake of my brethren and my friends~\sep\ I will speak peace concerning You.~\sep

  For the sake of the house of the Lord our God~\sep\ I will seek your good.~\sep

  \Psalm{122} I lift my eyes to You, the One who dwells in the heavens.~\sep

  As the eyes of servants are toward their masters,~\sep

  and as the eyes of a maidservant are toward her mistress,~\sep\ so our eyes are toward You, O Lord our God,~\sep\ until You have mercy on us.~\sep

  Have mercy on us, O Lord, have mercy on us,~\sep\ for we have heard much contempt.~\sep

  Our soul is exceedingly filled with (\translationoption{overcome by})~\sep\ the derision of mockers~\sep\ and the scorn of the proud.~\sep

  \Psalm{123} If the Lord had not stood up for us~\sep\ may Israel say,~\sep\ if the Lord had not stood up for us;~\sep

  when men rose up against us,~\sep\ they would have swallowed us alive,~\sep\ when their anger burned against us.~\sep

  They would have drowned us in the waters;~\sep\ the torrent would have passed over our soul.~\sep

  Many waters would have crossed over our soul.~\sep

  Blessed is the Lord, who did not give us as food for their teeth.~\sep

  Our soul was delivered like a bird from the snare of the hunter;~\sep\ the snare was broken, and we escaped.~\sep

  Our help is in the Name of the Lord,~\sep\ who made the heavens and the earth.~\sep

  \Psalm{124} Those who hope in the Lord are in mount Zion;~\sep\ they will not be shaken but will abide forever.~\sep

  As for Jerusalem, the mountains surround her;~\sep\ so the Lord surrounds His people~\sep\ from now and forever.~\sep

  For the rod of the wicked will not rest on the portion of the righteous;~\sep\ nor will the righteous stretch out their hands on iniquity.~\sep

  Do good, O Lord, to those who are good~\sep\ and to those who are upright in their heart.~\sep

  But as for those who pervert their paths,~\sep\ the Lord will drive them away with the workers of iniquity.~\sep\ But may peace be on Israel.

  Glory... --- Alleluia, alleluia, glory to You, O God, alleluia, alleluia, glory to You, O God, alleluia, alleluia, O Lord, have mercy on us.
\end{halfparskip}

\dd~Let us pray; peace be with us.

\cc~May the hidden power, O my Lord, of Your Divinity, and the wondrous help of Your Lordship, and the great aid of Your mercy strengthen the weakness of our feeble nature to lift up to You glory, honour, confession and worship at all times, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsunday{3. Sundays ``before''}{Marmita 31 (Ps 79--81)}

\begin{halfparskip}
  \Psalm{79} O God, the peoples have entered Your inheritance;~\sep\ and they have defiled Your holy temple.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). --- First verse.}

  They have made Jerusalem a desolation.~\sep\ They have given the corpses of Your servants as food to birds of heaven~\sep\ and the flesh of Your righteous ones to the beasts of the land.~\sep

  They have shed their blood like water around Jerusalem; and there is no one who buries them.~\sep

  We have become a reproach to our neighbours;~\sep\ and a derision and a scorn to those around us.~\sep

  How long are You angry, O Lord? Forever?~\sep\ and does Your anger burn like fire?~\sep

  Pour out Your anger on the peoples who have not acknowledged (\translationoption{known}) You~\sep\ and on the kingdoms that have not called on Your Name.~\sep

  For they have consumed Jacob; and have devastated his dwelling place.~\sep

  Do not remember against us our former sins;~\sep\ may Your mercies precede us,~\sep\ for we have been brought very low.~\sep

  Help us, O God our Saviour,~\sep\ for the honour of Your Name.~\sep

  Have mercy on us and deliver us from our sins for Your Name's sake,~\sep

  so that the peoples not say, Where is their God?~\sep\ May be known among the peoples in our sight~\sep\ the vengeance of the blood of Your servants that was shed.~\sep

  May the groaning of the prisoner come (\translationoption{enter}) before You;~\sep\ by the greatness of Your arm (\translationoption{power}) release the sons from death.~\sep

  Repay our neighbours sevenfold in their bosom~\sep\ the reproach with which they have reproached You, Lord.~\sep

  And we Your people and the sheep of Your pasture~\sep\ will give thanks to You forever;~\sep\ and we will relate Your wondrous deeds to generation after generation.

  \Psalm{80} Give ear (\translationoption{Give heed}), O shepherd of Israel;~\sep\ guide Joseph like sheep.~\sep

  Reveal Yourself, You who are seated above the cherubim.~\sep\ Before Ephraim, Benjamin, and Manasseh~\sep\ show Your strength and come for our salvation.~\sep

  Almighty God, turn us;~\sep\ and make Your face to shine, and we will be saved.~\sep

  Lord God almighty, how long will You be angry over the prayer of Your servant?~\sep

  You have made them eat bread with tears;~\sep

  You have made them drink with tears.~\sep\ You have made us a mocking to our neighbours;~\sep\ and our enemies deride us.~\sep

  Turn us, almighty God;~\sep\ make Your face to shine, and we will be saved.~\sep

  You brought forth a vine from Egypt;~\sep\ and You destroyed the peoples, and You planted it.~\sep

  You gave heed to it and planted its root;~\sep\ and the land was filled by it.~\sep

  The mountains were covered by its shadow;~\sep\ and its shoots reached over the cedars of God.~\sep

  It extended its tendrils as far as the sea,~\sep\ and its offshoots over the rivers.~\sep

  Why did You break down its hedge?~\sep\ and all those who pass along the way trample it.~\sep

  A boar of the woods has consumed it;~\sep\ and animals of the field have devoured it.~\sep

  Return, almighty God;~\sep\ and look down from the heavens and see,~\sep\ and visit this vine,~\sep

  the vine that Your right hand planted,~\sep\ and the son of man whom You strengthened for Yourself.~\sep

  Burn its shoots with fire;~\sep\ may they perish by the rebuke of Your face.~\sep

  May Your right hand be on the man,~\sep\ and on the son of man whom You strengthened for Yourself.~\sep

  We will not turn away from You;~\sep\ but quicken us, and we will call on Your Name.~\sep

  O Lord God almighty, turn us;~\sep\ make Your face to shine, and we will be saved.

  \Psalm{81} Praise God, who strengthens us;~\sep\ and cry out to the God of Jacob, alleluia.

  Take up the tambourines and citherns;~\sep\ the pleasant (instruments) with the harps, alleluia.

  Blow the horns at the beginning of months,~\sep\ and at the full moons on the feast days, alleluia.

  For it is a statute for Israel;~\sep\ and a law of the God of Jacob, alleluia.

  He ordained it as a testimony in Joseph,~\sep\ when he went forth against (\translationoption{to}) the land of Egypt, alleluia.

  He heard a language that he did not understand, alleluia.

  ``I removed the yoke from his shoulder;~\sep\ and I loosed his hands from bondage, alleluia.

  In distress he called to Me, and I delivered him;~\sep\ and I covered him with My glorious protection.~\sep\ and I tested him at the waters of controversy, alleluia.

  Listen, My people, and I will speak;~\sep\ Israel, I will testify to you,~\sep\ if you will listen to Me, alleluia.

  There shall be for you no foreign god;~\sep\ nor shall you worship another deity, alleluia.

  I am the Lord your God,~\sep\ who brought you out from the land of Egypt.~\sep\ Open your mouth, and I will fill it, alleluia.

  But My people did not listen to My voice;~\sep\ Israel was not persuaded by Me, alleluia.

  They walked in the desire of their heart~\sep\ and in their own (\translationliteral{of their soul}) belief (\translationoption{opinion}), alleluia.

  O that my people had listened to Me! and that Israel had walked in My ways, alleluia.

  I would have quickly destroyed their enemies;~\sep\ and I would have turned My hand against their foes (\translationoption{those who hate me})'', alleluia.

  Those who hate the Lord have denied (\translationoption{dealt falsely with}) Him;~\sep\ their tottering (\translationoption{impulse}) is forever, alleluia.

  He fed him from the best wheat;~\sep\ and he satisfied him with honey from the rock, alleluia.
\end{halfparskip}

\markedsunday{Sundays ``after''}{Marmita 51 (Pss 125-129)}

\begin{halfparskip}
  \Psalm{125} When the Lord was restoring the captivity of Zion,~\sep\ we were like those who are glad.~\sep

  \liturgicalhint{Alleluia, alleluia, (alleluia). --- First verse.}

  Then our mouth will be filled with laughter, and our tongue with praise.~\sep

  Then it will be said among the peoples,~\sep\ The Lord has done great things with them.~\sep

  The Lord has done great things for us; and we are glad.~\sep

  Turn back, O Lord, our captivity like the brooks of the south.~\sep

  Those who sow with tears will reap with gladness.~\sep

  The one who goes forth weeping,~\sep\ carrying seed,~\sep

  will surely come with gladness,~\sep\ bringing the sheaf.

  \Psalm{126} Unless the Lord builds the house,~\sep\ its builders labour in vain.~\sep

  Unless the Lord watches over the city,~\sep\ its watchers keep vigil in vain.~\sep

  Those who rise early and those who sit up late do so in vain;~\sep

  they eat their bread (\translationoption{food}) with grief.~\sep\ Thus He will give sleep to His beloved ones.~\sep

  For children are an inheritance of the Lord,~\sep\ the reward of the fruits of the womb.~\sep

  As an arrow in the hand of a warrior (\translationoption{mighty man}),~\sep\ so are the children of one's youth.~\sep

  Blessed is the man whose quiver is full of them;~\sep

  they will not be ashamed when they speak with the enemy in the gate.

  \Psalm{127} Blessed is everyone who fears the Lord~\sep\ and walks in His paths.~\sep

  When you eat the labour of your hands, son of man,~\sep\ you will be blessed, and it will be well for your soul.~\sep

  Your wife will be like a magnificent vine on the walls (\translationliteral{borders}) of your house;~\sep

  your children will be like an olive planting around your table.~\sep

  Thus will be blessed the man who fears the Lord.~\sep

  May the Lord bless you from Zion;~\sep\ may you see the prosperity of Jerusalem all the days of your life.~\sep

  May you see your children's children.~\sep\ Peace be on Israel.

  \Psalm{128} My oppressors have been numerous from my youth,~\sep\ may Israel say.~\sep\ My oppressors have been numerous from my youth,~\sep

  but they have not overcome me.~\sep

  They beat me with scourges on my back; and they continued their humiliation.~\sep

  The Lord is righteous; and He cuts off the branches of the wicked.~\sep

  May they turn (\translationliteral{be turned}) backward, all those who hate Zion.~\sep

  May they be like grass on the roof,~\sep\ which when the wind blows on it dries up and withers.~\sep

  The reaper cannot fill his hand with it,~\sep\ nor can the one who binds sheaves (fill) his arm.~\sep

  Nor do those who pass by say,~\sep\ ``May the blessing of the Lord be on you;~\sep\ we bless you in the Name of the Lord''.

  \Psalm{129} From the depth I called out to You, O Lord; and You heard my voice, alleluia.

  May Your ears be attentive to the voice of my supplication, alleluia.

  If You were keeping in memory the sins, O Lord, who would be able to stand? alleluia.

  For forgiveness is from You, alleluia.

  I have hoped in the Lord; and my soul has waited for His word, alleluia.

  I waited for the Lord from the morning watch until the morning watch, alleluia.

  May Israel wait for the Lord,~\sep\ for mercies come (\translationoption{are}) from Him.~\sep\ and with Him there is plenteous redemption, alleluia.

  He will save Israel from all its iniquity, alleluia.

  Glory... ---~\sep\ Alleluia, alleluia.
\end{halfparskip}

\dd~Let us pray, peace be with us.

\cc~My Lord, for Your Nature, which is hidden, incomprehensible, and unlimited for the thinking and understanding of all creatures, glory is fitting and the sound of praise is right; worship is seemly and due from all whom You have created and formed in heaven and in earth, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Qalta}

\liturgicalhint{(Proper, see Hudra.)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\end{document}