\documentclass[12pt,twoside,a5paper]{article}

\usepackage{multicol}

\usepackage[main=english]{babel}
\usepackage{divine-office}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

% Version: 2024-07-13

\begin{document}

\title{Sapra~--- Ferial}
\author{}
\date{}
\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\dd~Let us pray; peace be with us.

\cc~My Lord, all creatures which You have created, which have been fashioned by Your beckon, who by Your will have come into being, come before You and kneel and worship, sing and praise, ask and entreat, pray and beseech, confess and glorify You, for You are the cause of (all) beings, and the breath and respiration of our lives, Lord of all, Father, Son, and Holy Spirit forever.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Psalms}

\begin{halfparskip}
  \Psalm{100} Praise the Lord, all the earth.~--- \qanona{Giver of light, Lord, we lift up glory to You.}~\sep

  Praise the Lord, all the earth.~\sep\ Serve the Lord with gladness;~\sep

  Enter before Him with a hymn.~\sep\ Know that He is the Lord our God.~\sep

  It is He who made us, and not we ourselves.~\sep\ We are His people and the sheep of His pasture.~\sep

  Enter His gates with thanksgiving and His courts with praise;~\sep\ give Him thanks and bless His Name.~\sep

  For the Lord is good, and His grace is forever;~\sep\ and His faithfulness is (\translationoption{lasts}) for generations.~\sep

  Glory to the Father, the Son, and the Holy Spirit. From eternity and forever, amen.

  \enrepeatfirstverse{Praise the Lord, all the earth.}~--- \qanona{Giver of light, Lord, we lift up glory to You.}
\end{halfparskip}

\dd~Let us pray, peace be with us.

\cc~To You, my Lord, we lift up glory and confession, and continually offer worship and thanksgiving to Your living and holy Name, for You are the Lord and Creator of all, Father, Son and Holy Spirit forever...

\begin{halfparskip}
  \Psalm{91} He who sits in the shelter (\translationoption{hiding-place}) of the Most High,~--- \qanona{Glorious is Your confidence, Christ, our Saviour.}~\sep

  He who sits in the shelter (\translationoption{hiding-place}) of the Most High~\sep\ and is glorified in the shadow of God,~\sep

  said to the Lord, You are my confidence;~\sep\ God is my refuge in whom I trust.~\sep

  For He will deliver you from the snare of scandal~\sep\ and from vain talk.~\sep

  On His pinions (\translationoption{feathers}) He will deliver you; and under His wings you will be sheltered;~\sep\ (like) armour His truth will surround you.~\sep

  Do not fear the terror of the night,~\sep\ nor the arrow that flies in the daytime,~\sep

  nor the word that walks in the darkness,~\sep\ nor the wind that devastates in the midday.~\sep

  Thousands will fall at your side;~\sep\ tens of thousands at your right hand.~\sep\ But they will not draw near to you.~\sep

  But with your eyes you alone will see;~\sep\ you will look on the punishment of the wicked.~\sep

  For You, O Lord, are my confidence;~\sep\ You have placed Your dwelling in the height.~\sep

  Misfortune (\translationoption{evil}) will not draw near to you;~\sep\ nor will disaster draw near to your dwelling.~\sep

  For He will appoint His angels over you,~\sep\ so that they may guard you in all your paths.~\sep

  Upon their arms they will carry you,~\sep\ so that you do not stumble with your foot.~\sep

  You will tread on the asp and the cockatrice;~\sep\ you will trample the lion and the dragon.~\sep

  Because he has sought me, I will deliver him and strengthen him.~\sep\ Because he has known My Name, He will call on Me, and I will answer him.~\sep

  I will be with him in distress;~\sep\ I will strengthen him and honour him.~\sep

  I will satisfy him with length of days; and I will show him My salvation.~\sep

  \enbriefglory\ \enbrieffrom\~--- \enrepeatfirstverse{He who sits in the shelter (\translationoption{hiding-place}) of the Most High.~\sep}

  \qanona{Glorious is Your confidence, Christ, our Saviour.}
\end{halfparskip}

\dd~Let us pray, peace be with us.

\cc~Glorious, my Lord, is the great confidence in Your Godhead, and high and exalted is the wondrous refuge of Your Majesty, and all who hope in You, trust in You, call Your holy Name, and beseech You at all seasons and times, will not be ashamed, Lord of all...

\begin{halfparskip}
  \Psalm{104,1--16} Bless the Lord, my soul!~--- \qanona{May the glory of the Lord be for ever.}~\sep

  Bless the Lord, my soul.~\sep\ O Lord my God, You are very great;~\sep\ You are clothed with brightness and honour.~\sep

  You are covered with light, as with a mantle;~\sep\ You have stretched out the heavens like a curtain.~\sep

  He has made His habitations on the waters;~\sep\ and has placed His chariot on the clouds,~\sep\ and He walks on the wings of the wind.~\sep

  He has made His angels spirit~\sep\ and His ministers a blazing fire.~\sep

  He has established the earth on its foundations,~\sep\ so that it can never totter.~\sep

  You covered it with the deep, as with a mantle;~\sep\ and the waters stood above the mountains.~\sep

  At Your rebuke they flee;~\sep\ and at the sound of Your thunders they are terrified.~\sep

  They ascended to the mountains and descended to the valleys (\translationoption{plains}),~\sep\ to the place that You established for them.~\sep

  You have set a boundary for them,~\sep\ so that they may not cross over and cover the earth.~\sep

  You sent springs in the valleys;~\sep\ they travel among the mountains.~\sep

  They give to drink to every animal of the field;~\sep\ and wild donkeys that thirst are satisfied.~\sep

  Birds of the heavens settle on them;~\sep\ and they utter (\translationliteral{give}) a sound among the mountains.~\sep

  He irrigates the mountains from His habitations;~\sep\ the earth is satisfied by the fruits of Your works.~\sep

  He makes grass grow for the cattle~\sep\ and green herb for the cultivation of the sons of men.~\sep

  so that they may bring forth bread (\translationoption{food}) from the earth:~\sep\ wine that gladdens the heart of man,~\sep

  oil to make him a cheerful countenance,~\sep\ bread (\translationoption{food}) that sustains the heart of man.~\sep

  The trees of the Lord are satisfied (\translationoption{well watered}).~\sep

  \liturgicalhint{No Glory; only qanona:} \emph{May the glory of the Lord be for ever.}
\end{halfparskip}

\begin{halfparskip}
  \Psalm{113} Praise, servants of the Lord.~--- \qanona{Glory to Him who created the light.}~\sep

  Praise, servants of the Lord;~\sep\ praise the Name of the Lord.~\sep

  Blessed be the Name of the Lord~\sep\ from always and for ever.~\sep

  From the rising of the sun to its setting~\sep\ the Name of the Lord is great.~\sep

  The Lord is high above all the peoples,~\sep\ and His honour is above the heavens.~\sep

  Who is like the Lord our God,~\sep\ who dwells on high and looks on the depths~\sep\ in heaven and on earth?~\sep

  He raises the needy from the dunghill,~\sep\ so that He might seat him with the princes of the people.~\sep

  He causes the barren woman to dwell at home (\translationoption{to keep house})~\sep\ as a joyful mother of children.~\sep

  \liturgicalhint{No Glory. No Qanona.~--- Then \Pss{93; 148--150; 117} without farcing (= interpolation) under one Glory.}
\end{halfparskip}

\begin{halfparskip}
  \Psalm{93} The Lord reigns and He is clothed in magnificence.~\sep\ The Lord is clothed in strength and is mighty.

  He has established the world, so that it will not totter.~\sep

  Your throne was established long ago;~\sep\ and You are from everlasting.~\sep

  The rivers are lifted up, O Lord,~\sep\ the rivers have lifted up their voice;~\sep\ the rivers are lifted up in purity.~\sep

  More glorious than the sound of many waters,~\sep\ the mighty storms of the sea,~\sep\ is the Lord in the heights.~\sep

  Your testimonies are very trustworthy;~\sep\ and holiness becomes Your house,~\sep\ O Lord, for length of days.~\sep
\end{halfparskip}

\begin{halfparskip}
  \Psalm{148} Praise the Lord from the heaven;~\sep\ praise Him in the heights.~\sep

  Praise Him, all His angels;~\sep\ praise Him, all His hosts.~\sep

  Praise Him, sun and moon;~\sep\ praise Him, all the stars and light.~\sep

  Praise Him, the highest heavens, (\translationliteral{the heavens of the heaven})~\sep\ the waters that are above the heavens. May they praise the Name of the Lord.~\sep

  For He spoke, and they came into being;~\sep\ He commanded, and they were created.~\sep

  He established them forever and ever;~\sep\ He gave a law, and it does not pass away.~\sep

  Praise the Lord from the earth,~\sep\ dragons and all depths,~\sep

  fire and hail, snow and frost,~\sep\ winds and hurricanes, the work of His word.~\sep

  mountains and all hills,~\sep\ fruit trees and all cedars,~\sep

  animals and all cattle,~\sep\ creeping things and birds and flying creatures,~\sep

  kings of the earth and all peoples,~\sep\ princes and all judges of the earth,~\sep

  young men and maidens,~\sep\ old men and youths. May they praise the Name of the Lord.~\sep

  For great is His Name alone,~\sep\ and His glory in the earth and in the heavens.~\sep

  He raises up the horn of His people, and the praise of all His righteous ones,~\sep\ the children of Israel, a people who are near to Him.~\sep
\end{halfparskip}

\begin{halfparskip}
  \Psalm{149} Sing to the Lord a new song,~\sep\ His praise in the assembly of the righteous.~\sep

  May Israel be glad in its Maker;~\sep\ may the children of Zion rejoice in their King.~\sep

  May they praise His Name with tambourines and timbrels;~\sep\ and may they sing to Him with harps.~\sep

  For the Lord takes pleasure in His people;~\sep\ and He gives salvation to the poor.~\sep

  May the righteous be strengthened in honour;~\sep\ and may they praise Him on their beds.~\sep

  May they exalt God with their throats,~\sep\ with a double-edged sword in their hands~\sep

  to exact (\translationoption{inflict}) vengeance from the peoples~\sep\ and reproof from the nations,~\sep

  to bind their kings with fetters~\sep\ and their nobles with chains of iron,~\sep

  in order to carry out against them the judgment that is written.~\sep\ Praise to all His righteous ones.~\sep
\end{halfparskip}

\begin{halfparskip}
  \Psalm{150} Praise the Lord in His sanctuary;~\sep\ praise Him in His mighty firmament (\translationoption{firmament of His strength}).~\sep

  Praise Him in His might;~\sep\ praise Him in the multitude of His greatness.~\sep

  Praise Him with the sound of the horn;~\sep\ praise Him with lyre and with harp.~\sep

  Praise Him with tambourines and timbrels;~\sep\ praise Him with pleasant (\translationoption{sweet sounding}) strings.~\sep

  Praise Him with loud cymbals;~\sep\ praise Him with voice and with shouting.~\sep

  May everything with breath praise the Lord.

  \Psalm{116} Praise the Lord, all you peoples;~\sep\ praise Him, all nations.

  For His graciousness to us is great (\translationoption{strong}); truly the Lord is forever.

  Glory...~--- Praise the Lord, all the earth.~--- \qanona{Giver of light, Lord, we lift up glory to You.}
\end{halfparskip}

\Slota{\dd~Let us pray, peace be with us.}

\cc~Heaven and earth and all that is in them, are bound to confess, worship and glorify You, my Lord, for all Your help and graces towards us, the greatness of which cannot be repaid, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Laku mara}

\begin{doublecols}
  \englishl \rr~\liturgicalhint{(for singing)} Lord of all, we worship You; Jesus Christ, we exalt You; You give life to our bodies; and salvation to our souls!

  \englishc{1} \rr~You, Lord of all, we confess. And You, Jesus Christ, we glorify. For You are the Quickener of our bodies. And You are the Saviour of our souls.
\end{doublecols}

\dd~In the morning You hear my voice, Lord, in the morning I prepare myself and appear before You.

\rr~Lord of all...

\cc~Glory... From eternity and forever, amen and amen.

\rr~Lord of all...

\Slota{\dd~Let us pray, peace be with us.}

\cc~You, who quicken our bodies in Your compassion, and answer (us), and save our souls by a pleasing beckon of Your will, O great King of glory, Being who are from everlasting, we confess, worship, and glorify, at all seasons and times, Lord of all...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{\Ps{51,1--18}}

\begin{halfparskip}
  Have mercy on me, God, according to Your grace;~\sep\ and according to the multitude of Your mercies, blot out my sins.~--- \qanona{Have mercy upon me, Lord.}~--- \liturgicalhint{Repeat first verse.}

  Wash me thoroughly from my iniquity,~\sep\ and cleanse me from my sins.~\sep

  For I acknowledge my transgressions,~\sep\ and my sins are always before me.~\sep

  Against You alone have I sinned,~\sep\ and done evil deeds before You;~\sep

  for You shall be justified in Your saying,~\sep\ and blameless in Your judgments.~\sep

  For I was conceived in iniquity,~\sep\ and in sin my mother conceived me.~\sep

  But You delight in truth,~\sep\ and have made known to me the secret things of Your wisdom.~\sep

  Sprinkle me with hyssop, and I will be clean;~\sep\ purify me with it, and I will be whiter than snow.~\sep

  Satisfy me with Your joy and Your gladness,~\sep\ and my humbled bones will rejoice.~\sep

  Turn away Your face from my sins,~\sep\ and blot out all my iniquities.~\sep

  Create in me a clean heart, God,~\sep\ and renew a steadfast spirit within me.~\sep

  Cast me not away from Your presence,~\sep\ and take not Your holy spirit from me.~\sep

  But restore to me Your joy and Your salvation,~\sep\ and let Your glorious spirit sustain me;~\sep

  That I may teach the wicked Your way,~\sep\ and sinners may be converted unto You.~\sep

  Deliver me from blood-guilt, my God, God of my justification,~\sep\ and my tongue will praise Your righteousness.~\sep

  Lord, open for me my lips,~\sep\ and my mouth will recite (\translationoption{bring forth}) Your praises.~\sep

  For You do not take pleasure in sacrifices,~\sep\ nor are You appeased (\translationoption{pleased}) by whole burnt offerings.~\sep

  The sacrifices of God are a humble spirit;~\sep\ a contrite heart God will not despise.

  Glory...~--- \qanona{Have mercy upon me, Lord.} Have mercy on me, God, according to Your grace; and according to the multitude of Your mercies blot out my sins. Christ the King, have mercy upon me.
\end{halfparskip}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Tesbohta}

\liturgicalOption{I. For recital:}

\begin{halfparskip}
  To You be glory, our God, and to You be praise, our Maker.

  We bless You, who formed us, compassionate Lord; merciful God, pitiful Creator, Saviour, who preserves us.

  Who helps and aids us, we worship You, Lord; for You are long-suffering and Your loving-kindness is great.

  Merciful One, pity us and have mercy upon us; turn to us in Your abundant mercy.

  Our Hope and strong Refuge; Lord our God, make Your face to shine, and we shall be saved.

  You who receive the penitent in mercy, receive our prayer and our service.

  You who hear the voice of Your worshippers, let our request come before You, and have mercy upon us.

  You who forgive the debts of mortals in Your compassion forgive us our debts and sins in Your compassion.

  You who pardon the sins of the sons of men in Your loving-kindness, pardon our many sins, and have mercy upon us.

  Good Hope of the sons of men, give us peace and tranquillity, that we may confess Your Trinity, Lord of all, for ever, amen. \liturgicalhint{(3x)}
\end{halfparskip}

\liturgicalOption{II. For singing:}

\begin{doublecolsong}
  \textsizexi

  \songlinein{In Your presence, Lord of all,}

  \songline{we stand in humility}

  \songline{singing hymns of praise to You.}

  \songline{In Your mercy, Lord our God,}

  \songline{we will bless Your love and grace}

  \songline{Who are our own Creator.}

  \songindentv\songlinein{Lord our Saviour, with your hands}

  \songline{hold us safe at all moments;}

  \songline{protect and guide at all times.}

  \songline{You who are God, Lord of all,}

  \songline{have mercy on Your servants}

  \songline{who praise Your Divinity.}

  \songindentv\songlinein{Lord Jesus Christ, port of peace,}

  \songline{pour forth in floods Your mercy,}

  \songline{look upon us Your children;}

  \songline{let Your face shine on us, Lord,}

  \songline{illumine these humble ones,}

  \songline{that we may find salvation.}

  \songindentv\songlinein{Lord of mercy who receives}

  \songline{every sinner who repents,}

  \songline{You are truly our refuge.}

  \songline{Hear our prayers, kindly Lord;}

  \songline{let the cries of worshippers}

  \songline{rise before You like incense.}

  \songindentv\songlinein{Lord who free all from their debts,}

  \songline{in Your mercy, loose our bonds}

  \songline{and remit our trespasses.}

  \songline{You who pardon in mercy,}

  \songline{forgive us each all our sins}

  \songline{though so countless in Your sight.}

  \songindentv\songlinein{You are the true hope of all:}

  \songline{Grant peace and prosperity}

  \songline{to Your children in the world,}

  \songline{that we all may sing and bless}

  \songline{the Name of Your Trinity,}

  \songline{O Lord of all, for ever.}
\end{doublecolsong}

\Slota{\dd~Let us pray, peace be with us.}

\cc~Christ, the good Hope of the sons of men, and the peaceful Haven which give peace to the creatures; my Lord, grant us Your peace and tranquillity, that therein we may confess and glorify You without ceasing, all the days of our life, Lord of all forever. Amen.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Trisagion and Our Father}

\dd~Lift up your voices, and glorify the living God, all you people.

\begin{doublecols}
  \englishl \rr~Holy, holy, holy God\\
  Holy, holy, mighty One\\
  Holy and immortal God\\
  Lord, have mercy on us all!

  \englishc{1} \rr~Holy God, Holy Mighty, Holy Immortal, have mercy upon us.
\end{doublecols}

\begin{halfparskip}
  \rr~Let us pray, peace be with us.

  \liturgicalhint{If the Martyrs' Anthem is not said, the Lord's Prayer is omitted, and the anthem ``At the sunrise'' is immediately begun (see page \pageref{at-the-sunrise}).}

  Our Father in heaven, hallowed be Your Name; Your kingdom come. Holy, holy, holy are You. Our Father in heaven, heaven and earth are full of the greatness of Your glory; angels and men cry out to You: Holy, holy, holy are You.

  Our Father in heaven, hallowed be Your Name; Your kingdom come, Your will be done on earth as it is in heaven. Give us this day the bread we need; and forgive us our offences and sins as we have forgiven those who have offended us. Lead us not into temptation, but deliver us from the evil one. For Yours is the kingdom, the power and the glory for ever and ever. Amen.

  \cc~Glory be to the Father and to the Son and to the Holy Spirit.

  From eternity and forever. Amen and amen. Our Father in heaven, hallowed be Your Name; Your kingdom come. Holy, holy, holy are You. Our Father in heaven, heaven and earth are full of the greatness of Your glory; angels and men cry out to You: Holy, holy, holy are You.

  \cc~Compassionate One, whose Name is holy, o good and just One, whose grace and mercy are overflowing from everlasting, pour forth, our Lord and our God, the sweet tenderness of Your love on the souls of Your worshippers, who call on You and beseech You, at all seasons and times, Lord of all...

  \cc~Bless, my Lord, Your servants in Your loving-kindness and keep Your worshippers from all harm in the overflowing abundance of Your mercy; feed, lead, help, and protect our lives beneath the wings of the care of Your love; save, rescue, and deliver our bodies and our souls from the evil one and his hosts, at all seasons and times, Lord of all...
\end{halfparskip}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Martyrs' anthems}

\markedday{Mondays}

\begin{halfparskip}
  O Lord, in the morning You hear my voice (\translationoption{call}).~--- At morning time, when are opened the doors on high to prayer, receive, our Lord, our request, answer our petitions in Your mercy, and grant hope and salvation to the souls who wait for You.

  In the morning I prepare myself and appear before You.~--- The morning of our Lord has come, and suddenly our Saviour shines forth, and gives a reward to the just. Blessed is he who was zealous, and has worked in the vineyard of Christ, and has received his reward abundantly.

  Cause me to hear Your grace in the morning.~--- Make us worthy, our Lord, of that morning, when heaven and earth rejoice, and the just rejoice in their labours, and the righteous in their recompense, and the martyrs also in their crowns, and the nations in the forgiveness of their debts.

  In the morning my prayer will come before You.~--- At morning time Moses the chosen prayed on the top of the mount. There God answered him, and gave him the staff of help, that he might descend and save Israel from the subjection of the Egyptians.

  To declare Your graciousness in the morning.~--- At morning time Moses saw a terrible vision on the top of the mount. God was solemnly carried, cherubim bearing His throne, seraphim singing Holy in His honour, and angels serving Him.

  Sing to the Lord with the harps and with the sound of music.~--- At morning time the boy David sang with his harp songs of the Holy Spirit, and the animals assembled, and came to the sweet sound of his words when he sang alleluia.

  Let us praise Him with songs.~--- At morning time the martyrs sang a new song in the judgment hall: ``We will not deny Christ, who for us endured suffering, death and the Cross, and saved our race from error''.

  God sent His grace and His truth.~--- At morning time the dove was sent forth by perfect Noah because of the waters of the flood. And it turned and came to the ark, bearing an olive leaf, a sign of peace and tranquillity.

  You have given a sign to those who fear You.~--- At morning time a sign of peace and tranquillity was seen: The bow of our Lord in the clouds; and the mouth of the Creator swore: ``I will not cause a flood henceforward and for ever''.

  In the eyes (\translationoption{sight}) of the peoples.~--- In the morning the martyrs in the judgment-hall raised and lifted up their voices: ``We do not worship idols, we worship the one God, and confess His Son Jesus. He will deliver us from the fire''.

  Be strong, and let your heart be strengthened.~--- In the morning the martyrs in the judgment-hall comfort one another and say: ``The Lord is our helper, we will not fear. What does man unto us, whose crown and power perish? But the kingdom of the Lord is for ever''.

  Hear this, all you peoples.~--- At morning time was heard a voice which said to the athletes: ``Take up your crosses and follow Me, for I am your Lord and your Master, your God and Strengthener; and you are My disciples''.

  He carries out justice (\translationoption{judgment}) for the oppressed.~--- In the morning the faithful Shmuni offered a prayer to God: ``O God, execute judgment for me against the impious Antiochus, who kills my sons as lambs and threatens (\translationoption{assaults}) me as a lion''.

  And your prayers be on all of us.~--- Gadai and Maccabaeus with Tarsai, Hebron and Hebson with Bakos, and the seventh son, Jonadab, and their teacher Eleazar, and their mother, the faithful Shmuni, may their prayer be a wall to us.

  The camp of the angels of the Lord.~--- At morning time the martyrs went forth, ranks and ranks of saints, from the prison, saying: ``Hasten, king, to cut off our heads, that we may depart and go up to heaven, to our Lord, who waits for us''.

  For He has done a marvellous thing.~--- That which was wrought from one end of the world to the other was a marvel in the beloved boy Cyriac at the age of three years. Grace overshadowed him, and he was strengthened in the faith.

  Just and upright.~--- May the prayer of the martyr George be to us a high rampart, that the arrows of the evil one may astray, and that he does not pierce us with his spearhead, he who hates our race at all times, and ever wishes for our destruction.

  Come, hear! I will relate to You.~--- Hear and wonder, sagacious ones. A virgin has borne in Bethlehem the Light of the whole world. And lo, her memorial is celebrated in the four quarters of creation, and in heaven amongst the angels.

  Glory...~--- Glory to You at morning time from those in the height and below, O Son who sit at the right hand, for with the sound of the horn are awakened old men, young men and children, whom You have created to praise You.

  From eternity and forever.~--- Blessed is he who has traded, as you, our father, have traded. He has collected spiritual riches, and has filled his ship with all good things, and has steered and come to the haven, to the meeting place of all the just.

  May all the people say, Amen and amen.~--- Our Lord, Your kingdom come; Your will be done on earth as it is in heaven. Give us the bread of our need, and lead us not into temptation, but deliver us from the evil one.

  May the help which accompanied Moses, the firstborn of the prophets, and who divided the sea before the people, accompany Your worshippers, who have come to ask pardon from the treasury of Your kindness.
\end{halfparskip}

\markedday{Tuesdays}

\begin{halfparskip}
  Praise the Lord, O righteous ones.~--- Martyrs, you were merchants. And lo, your treasure is in heaven. You have bought the pearl with the blood that your necks poured forth.

  Praise is becoming for the upright.~--- The martyrs were raised and reached the heavenly Jerusalem, and bought with the blood of their necks the country for which they were lying in wait.

  They travelled from people to people.~--- You departed from one place to another, but departed not from your Lord. And in every place where you entered, you placed the treasure of life.

  The Lord will dwell in it forever.~--- In the place where your bones are laid dwell peace and tranquillity. And the hosts of the watchers on high keep it and its inhabitants.

  Blessed are the people for whom this is the case.~--- Blessed is the place where are laid your bones as treasures. Whenever the light of the sun gets down, the light from your bones has shone forth.

  Light has risen in the darkness for the upright.~--- A light has shone forth from your bones, and led away captive my mind for (\translationliteral{after}) you. I went forth to see your deaths and I wondered about your crowns.

  They are more to be desired than gold and precious stones.~--- Your bones are pearls, and your bodies are sweet incense. The blood that flowed from your necks became grace for the whole creation.

  Those who are in the graves will relate Your grace.~--- Lo, in the Church are your bones; and you have the keys of the heights [= heaven]. Blessed ones, open your treasures, and give help to the needy.

  Like a city that is surrounded by a wall.~--- Like beams in a building, the martyrs girded creation. And when the earth was about to fall, they supported it by the power of their bones.

  I gathered my strength (\translationoption{was strengthened}), but I could not prevail (\translationoption{was unable}).~--- The sea cannot make a breach in the wall which your love surrounds and Satan cannot rule in the land where your bones are laid.

  For Your sake we are killed all day.~--- The Lord, whose martyrs were killed for the glory of His Majesty, in the day when His Cross shines forth, will place them at His right hand.

  For there the Lord has commanded blessing, even life forever and ever.~--- Where the martyrs were killed, and their limbs cut off, there the Holy Spirit came down, and made peace amidst desolation.

  Come! Let us kneel down and worship Him.~--- Before your graves let us reverence the power hidden in your bones; and we, who have seen your deaths on earth, will see your crowns in heaven.

  Those who go down to the sea in ships.~--- Martyrs, you were merchants, and journeyed by sea and land. You have brought peace to the sea by (your) tears, and to the land by your prayers.

  From the depth I called out to You, O Lord; and You heard my voice.~--- Out of the deep did the martyrs cry to You. And Your loving-kindness heard them. Out of the deep we call to You: Come, help us and save us.

  I will bless the Lord at all times.~--- Blessed be the Lord, whose servants have fallen asleep, and He exalts their memorials. They sleep, and are in peace and at rest; and from them help springs forth.

  How good and how fair it is.~--- How fair is the crown on your heads, woven by faith, and adorned by the Holy Spirit with precious and beautiful stones.

  Like a thirsty and buffeted land, that is in need of water.~--- As a land that thirsts for water, and daily longs for irrigation, so thirsted the martyrs for the love of their dear Lord.

  To new ones that were just made.~--- Come in peace, you new bridegrooms, sons of the mystery of baptism, whom grace made to grow in the Name of the Trinity.

  I was speaking peace.~--- Peace be to your bones, and rest to your limbs, for you were like your Lord, by whose death sinners have obtained life.

  Give Him thanks and bless His Name.~--- Confess the truth and do not slacken. Behold the swords and tremble not, for you have put on hidden armour, forged by the Holy Spirit.

  The small and the great.~--- The martyr said to his fellow, feeling compassion for his companion: ``Come, let us die for Jesus, and inherit eternal life''.

  The Lord is faithful in His words.~--- The martyrs heard this word which our Lord spoke, and were strengthened by it: ``He who dies for Me has everlasting life''.

  They are more to be desired than gold and precious stones.~--- The martyrs, who were killed for Jesus, are like precious stones, and valuable pearls.

  How good and how fair it is.~--- How beautiful is that ship, which bore mar Cyriac. Companies of cherubim accompany it, and seraphim singing ``Holy''.

  From the rising of the sun to its setting.~--- Lo, in all quarters is celebrated the glorious day of the memorial of the famous martyr Hormizd. Blessed is He who crowned him with victory.

  He heals the broken-hearted; He binds up their pains.~--- You were a healer, mar George, without medicines and roots. And every one who takes refuge in you is helped by your prayers.

  I will seek your good.~--- Beseech, you who are full of grace, the Child who shone forth from your bosom, that He make His peace and tranquillity to dwell in His Church, bought by His blood.

  Glory...~--- Peace to you who have conquered; peace to you who have been victorious. Peace to you who have become partakers in the sufferings of the Son.

  From eternity and forever.~--- Your truth is great, our father [patron saint]. Your beatitude is exalted and glorious. Intercede for us all, that we may share in your bliss.

  May all the people say, Amen and amen.~--- You who hear the prayers of Your servants, answer the petitions of Your worshippers. Hear our prayer and the sound of our request, and answer our petitions in Your mercy.

  You whose right hand is stretched out, whose door is open to the penitent, open the door to our prayer, and let our request come before You.

  You, around whose throne stand the fiery ones and the companies of the watchers, withdraw from us the deceiver, and rebuke the evil one, and let him be silenced.

  You who are true and who lie not, who have promised, ``Knock and I open'', be pleased with the prayer of Your worshippers, and hear the petitions of our souls.
\end{halfparskip}

\markedday{Wednesdays}

\begin{halfparskip}
  Your servants will give You thanks, O Lord.~--- Holy martyrs and teachers of faith, pray that there may be peace in creation. Let wars be brought to nought and contentions cease among us. And may the Church sing praise by the mouth of her children.

  Your righteous ones will praise You.~--- May the holy martyrs who confessed You in their afflictions, and propitiated You by the blood which their necks poured forth, be intercessors for the sinners with You, O our Lord, that in the day of judgment You may forgive their trespasses.

  They will speak of the glory of Your kingdom.~--- The martyrs saw the glory of Your kingdom in their minds, when they were being killed by their persecutors. And they joyfully endured dire tortures in their bodies. Our Lord Jesus received their spirits.

  Their blood is precious in His sight.~--- Let us diligently honour with songs of the Holy Spirit the bones of the martyrs who endured afflictions, that we may find help in the day of the recompense of their labours from God's merciful goodness.

  Glorious is the Lord in the heights.~--- On high are your crowns, and in the world are your assemblies, martyrs, preachers of Christ the King. Lo, on high and in the deep your feasts are celebrated, O sowers of peace in the four quarters.

  Seek the Lord and be strong.~--- Martyrs, seek mercy from the merciful God, that He may make His peace to dwell in the four quarters. And when our Lord is revealed and the clouds bear up your bodies, pray that with you we may inherit the kingdom.

  Lord, in the morning You hear my voice.~--- In the morning the martyrs cried in the judgment hall before the persecutors. ``We will not deny the heavenly Bridegroom, for He it is who delivers us from the hands of the ungodly, and clothes our bodies with glory in His kingdom''.

  In the morning I prepare myself and appear before You.~--- In the morning, when Christ the King, who crowns you, shines forth, O prophets, apostles, martyrs, and confessors, and (when He) places on your heads crowns of glory which fade not, pray that with you we may inherit the kingdom.

  Light has risen in the darkness for the upright.~--- O Martyrs, sons of light, who hated the world and its goods, and desired and loved the heavenly Bridegroom, beseech and intercede for us your Lord whom you loved, that He have pity on us and save our souls.

  Blessed are those who are without blemish in the way.~--- Blessed are you, martyrs who conquered Satan, and trod under foot the sting of death and sin. May your prayer be to us a high wall and house of refuge from the wiles of the evil one, the enemy.

  And who walk in the law of the Lord.~--- A blessing that passes not away and a kingdom that does not end has our Saviour promised to the friends whom He has chosen. And instead of afflictions, tortures and lacerations which they endured, He has given them to inherit life without end.

  Righteous and upright.~--- Small was in his youth the brilliant martyr Cyriac, and he astonished with his words the king and his servants, and then received the sword of the impious king. By it he was crowned for our Saviour's sake.

  He is fairer in his appearance than the (other) children of men.~--- A beautiful shoot, and an olive tree adorned with branches was mar George, the friend of Christ, who by the abasement and love which he had towards his Lord, stretched forth his shoulder and bore the Cross.

  They will lead her virgin companions after her.~--- Holy virgin Mary, blessed mother, all nations call you blessed, for you were worthy to carry in your bosom Emmanuel, whom the prophets described beforehand in their mysteries.

  Glory...~--- Glory to the Power which dwells in the bones of the saints. Its voice thunders in the four quarters. They are laid in the churches; help wells forth from them, and truth is revealed by the power of their words.

  From eternity and forever.~--- You who loved purity all the days of your life, chosen one, you were blameless for Christ. And because of your way of life, lo, your soul dwells in the tabernacle of the saints among the angels.

  May all the people say, Amen and amen.~--- Give us, our Life-giver, that which You know will help us. For we know not what we [should] ask. But one thing we beg for, that Your will may be accomplished in us, and that Your mercy may intercede for us all.

  Christ, who promised, ``To everyone who cries, I answer. And to him who knocks at My door, I open.'' Let not Your mercy and Your loving-kindness be far from Your worshippers, who take refuge in You, that they may receive help.
\end{halfparskip}

\markedday{Thursdays}

\begin{halfparskip}
  For He speaks peace over His people and with His righteous ones.~--- Peace to the martyrs, honour to their bones, glory to their Lord, and help to us by their prayer.

  In Your light we see light.~--- The martyrs are in the light, and the apostles in the bride-chamber of light; and lo, they sing glory to the Being who dwells in the excellent light.

  Speak reverently with the voice.~--- The martyrs, when they enter into the kingdom, say: ``Glory to You, our Lord, for we, who trust in You, are not confounded.''

  May none of those who hope in You be ashamed.~--- The martyrs say, ``We have taken refuge in Your Cross, merciful Lord, slacken not (Your) hands upon us.''

  He made a breach in the sea and made them cross over.~--- Martyrs, who passed over the bridge of fire to the high place, pray that we sink not in the troubled sea of sins.

  Their blood is precious in his sight.~--- Christ, who accepted the blood of the martyrs on the day when they were killed, accept our petition, and open the door to our prayer.

  Glorious is the Lord in the heights.~--- Above on high, and in the boundless treasure are laid up the labours of the martyrs, the friends of Christ.

  Your throne was established long ago.~--- Son, whose royal throne is established on high, by the prayer of Your saints, sustain the earth, which our trespasses have overthrown.

  Sing to the Lord with the harps and with the sound of music.~--- With the sound of alleluias and songs of the Holy Spirit, the saints sing, and the companies of legions [= devils] tremble.

  O Lord, in the morning You hear my voice.~--- Every morning the martyrs stand in prayer, and ask for help for the souls of sinners.

  In the morning I prepare myself and appear before You.~--- In the last morning, when the holy martyrs receive the reward of their labours, in that hour, have pity on us, our Lord.

  Offer Him sacrifices of praise.~--- May the prayers of the martyrs, who were pure offerings to their Lord, be unto us a wall by night and day.

  Blessed are those who are without blemish in the way.~--- Blessed are the martyrs when they hear that word: ``Come, enter, you blessed of My Father, and inherit endless life''.

  And who walk in the law of the Lord.~--- Blessed are you, whom your Lord has placed in a treasury, and on the altar is your memorial for ever.

  Do good, O Lord, to those who are good.~--- Blessed are the athletes, who represent the beauty of their persons in the heavenly way of life, so that it does not remain on earth and get disfigured.

  And to those who are upright in their heart.~--- Blessed are the martyrs at that time when are divided the good from the evil ones; and they inherit the kingdom.

  The camp of the angels of the Lord.~--- You of the family of Gregory [of Nazianzen], Basil and John [Chrysostom], pray that enter not into the Church the darkness of the heretics.

  And your prayers be on all of us.~--- You of the family of mar Ephrem and all your teachers of the truth [\emph{Hudra:} mar Ephrem, mar Narsai, and mar Abraham], may your prayers be to us a wall by night and day.

  Where there was no(thing to) fear.~--- All the holy martyrs have joy in the bride-chamber of light, on the right hand of Christ.

  How good and how fair it is.~--- How fitting, and how desirable is the glory with which the righteous are clothed in the day of the resurrection through the labours of their life.

  In the heavens and on earth.~--- Your body is with us, and your soul is in heaven, martyr George, beg for mercy from the Lord.

  Light has risen in the darkness for the upright.~--- Christ, who shone forth from blessed Mary, by the prayer of her who bore You, keep Your Church from harm.

  Glory...~--- Peace to the Church, and tranquillity to all who dwell in her; and praise to the Power which strengthened the martyrs in the contest.

  From eternity and forever.~--- May the prayer of our father, which was a pure offering to his Lord, be unto us a wall by night and day.

  May all the people say, Amen and amen.~--- Let not, my Lord, the prayer of Your servants be in vain, but for propitiation and forgiveness of sins.

  I entered into Your house and worshipped before Your judgment seat, merciful Lord, forgive me my debts and sins.

  This house is a type of heaven. In love without guile, let us enter into it continually.
\end{halfparskip}

\markedday{Fridays}

\begin{halfparskip}
  I will bless the Lord at all times.~--- Blessed is the hidden power which dwelt in the bones of the martyrs. For they are laid in their graves, and drive away devils from the world. By their teaching they brought to nought all the error of carved (idols). They secretly visit the creation and teach (men) to worship You. For You alone are the Lord.

  At all times His praises will be in my mouth.~--- Blessed is Christ, who clothed his twelve with strong armour. And they went forth to the four quarters and preached in the world His doctrine. And they destroyed the power of the enemy by the sword of the Spirit and converted the nations from error to the truth of their faith.

  Great is our Lord, and greatly to be praised.~--- Great is the truth of the holy martyrs, which is preached in all regions. Whom neither fire nor sword, nor the wrath of the pagan king, nor any other thing moved nor terrified. For God alone did they confess, and despised all tortures.

  The Lord is faithful in His words.~--- The just martyrs, who conquered and triumphed, were tested on the faith.

  For they knew that was true that word which our Lord said: ``Everyone who confesses Me before judges and rulers, I will cause to inherit the kingdom on high which is in heaven and I will confess him before the Father''.

  Look to Him and have hope in Him.~--- With the eye of the Spirit, the sons of the right hand looked, and verily saw the country which is above within the veil of the Holy of Holies. And they longed for it, and ran towards it by the way of the tortures of their persecutors. And, as on a bridge, they crossed the sea of the world by the Cross to the high place, which is their dwelling.

  Cause Your right hand, O Lord, to rest upon me.~--- May the right hand of our Lord, who calls all to the kingdom, bless our assembly, that we may keep vigil on the memorial of the martyrs. And that we also may be helped by their prayers, that they may be to our race a high wall in the last day, when the heights and the depths are amazed, and the will of the Creator shall reign.

  The camp of the angels of the Lord.~--- My mind wondered at the blessed company of athletes, the famous martyrs, how they despised and scorned this world and its desires. They looked at and saw with a bright eye the glorious brightness of the Pearl, which is at the head of the tree, and longed to seize it.

  And your prayers be on all of us.~--- By your prayers give help to the needy, O famous martyrs, for you have filled the earth with the good seed of your victories (\translationoption{heroic deeds}). And you have dug up the thorn-bush which the evil one sowed with his tares, and you sowed therein good seed, athletes, servants of truth.

  How good and how fair it is.~--- The athletes saw a pearl without flaw on the top of Golgotha, and longed diligently to meet it, and bought it with their own blood. And they endured sufferings and dire tortures for it. And lo, they have received the reward of their labours: endless felicity.

  Lifted up in all the earth.~--- High and exalted is the glorious and honoured bride-chamber which Christ the King has prepared for His athletes, who loved Him, believed and confessed His Name. And lo, their feasts are celebrated in the four quarters, while they cry to Him with the sound of alleluias, and say: ``Blessed is He who has exalted His saints''.

  Blessed are those who are without blemish in the way.~--- On the foundation of the twelve and the seventy apostles the diligent ones built: Mar Augin and mar Salita; and mar Milis with mar Babai, and the pious mar James of Soba (Nisibis), and mar Acha and mar John, may their prayer be a wall to us.

  I will bless the Lord at all times.~--- Blessed is Christ who strengthened the martyr mar Cyriac in his contest to stand against the persecutors, and he waxed strong and conquered the kings who contended with him. And by the strength of the Spirit he acted manfully, and endured stripes, and became a sacrifice for his Lord, and received the crown of victory.

  Exult, sing and offer praise.~--- With rejoicing the noble martyr mar George gave up his soul to all afflictions, for he heard Paul saying that the sufferings of this time are not to be compared to that glory which the just inherit at the last day. And upon him who here labours and acts zealously, the inheritance is bestowed.

  They will lead her virgin companions after her.~--- Virgin of the ages, Mary the blessed mother, beseech your Son to cause His peace to dwell in this temple, in which the day of your memorial is celebrated; and may He overshadow it with the right hand of His mercy as the house of Abram, so that all who are afflicted and distressed from Him may receive help.

  Glory...~--- Glory to You, good and kind Lord, by whose power the faithful ones triumphed, and by whose help they despised the threats of their persecutors. They destroyed the power of the enemy who boasted, for he saw that the martyrs did not weaken in their true hope.

  From eternity and forever.~--- May the prayer of our glorious father, who from his childhood pleased God like Samuel; who in humility was like Moses, the firstborn of prophecy; in zeal like Elijah; in love of his Lord like Abraham, be a wall to us from the wiles of the crafty one.

  May all the people say, Amen and amen.~--- We confess Your Godhead, good Lord. For by Your goodness You keep Your worshippers from harm; and, my Lord, You have opened Your door to them who repent and ask for mercy; and You stay not the great treasure of Your gift from Your flock, You, worshipped by all.
\end{halfparskip}

\markedday{Saturdays}

\begin{halfparskip}
  Lord, in the morning You hear my voice.~--- In the morning the martyrs were called to be killed, and the right hand of our Lord weaves their crowns.

  In the morning I prepare myself and appear before You.~--- In the morning the martyrs run to the contest, to receive a double reward of their labours.

  Cause me to hear Your grace in the morning.~--- In the morning our Lord called to the martyrs: ``Come, take the reward, zealous workmen''.

  Look to Him and have hope in Him.~--- In the morning the martyrs saw the Holy Spirit, weaving crowns and placing (them) on their heads.

  Light has risen in the darkness for the upright.~--- In the morning the martyrs shine forth like the sun, and go forth to meet the heavenly Bridegroom.

  His Name is holy and awesome.~--- In the morning the martyrs cry with the watchers: ``Holy, holy, holy God''.

  They will bring gifts to Him.~--- In the morning the martyrs make offerings: the prayer of their mouth and the blood of their necks.

  The offering of my hands like the evening offering.~--- In the morning the martyrs offer censers in their hands, and offerings to the King of kings.

  As the fragrance of sweet incense, and an acceptable savour before God.~--- In the morning the martyrs set forth incense to Him, whose chariot is harnessed with fire.

  Be strong, and let your heart be strengthened.~--- In the morning the martyrs encourage one another: ``Come, let us deride the impious king''.

  Before all the people.~--- In the morning the martyrs cry in the judgment-hall: ``We will not deny the heavenly Bridegroom''.

  He rebuked kings for their sake.~--- In the morning the martyrs stood in the judgment-hall, and refuted the kings and confounded the impious men.

  For Your sake we are killed all day.~--- You who were killed like Me, and persecuted for My sake, I and you (belong) to the Jerusalem above.

  We are considered as sheep for the slaughter.~--- The martyrs went to (\translationoption{were destined for}) death, and their Lord to (\translationoption{for}) the Cross; and the Holy Spirit weaves their crowns.

  Your throne was established long ago.~--- The throne is prepared and the King's Son sits; and the martyrs enter and receive their crowns.

  The small and the great.~--- In the morning the martyrs run to the Cross to drink the wine pressed out by the spear;

  The Lord will give strength to His people.~--- Grant the strength which strengthened the martyrs in the contest to the assembly of Your worshippers, our Lord.

  Open for me the gates of righteousness.~--- In the morning the King's Son opens His treasure, and the martyrs enter and receive their crowns.

  So that I may enter them and give thanks to the Lord.~--- In the morning our Lord opened paradise, and the robber entered and inherited the kingdom.

  When God was arising in order to judge~--- In the morning our Lord arises in the judgment, and calls to the martyrs, ``Come, take your crowns''.

  How good and how fair it is.~--- How fitting was it for the boy Cyriac when he rebuked the impious king.

  As a joyful mother of children.~--- In the morning Shmuni encourages her sons: ``Fear not him who kills men''.

  You will be blessed, and it will be well for your soul~--- Blessed are you, martyr George, who with the angels dwells in heaven.

  May mount Zion rejoice, and may the daughters of Judah be glad.~--- Let the faithful Church rejoice and exult in the memorial of Mary, the mother of God [\emph{Hudra:} mother of Christ].

  Glory...~--- Peace to the martyrs, and glory to their Lord; and to us, by their prayer, mercy and compassion.

  From eternity and forever.~--- Our glorious father, the friend of Christ, beg of your Lord mercy for us all.

  May all the people say, Amen and amen.~--- In the morning let us praise, and in the morning let us glorify the Lord of the morning, the true Light.

  In the morning let the prayer of Your worshippers enter before Your judgment-seat, merciful Lord.

  At morning time let us praise Your loving-kindness, for You have ordered the light for the joy of creation.
\end{halfparskip}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Fixed doxology}

\liturgicalOption{For singing (Mariam Btulta~/ Pulariprabhayil):}

\begin{doublecolsong}
  \songlinein{In the bright and dawning light}

  \songline{we are singing with delight}

  \songline{our hymns and praises to You,}

  \songline{Saviour of all creation.}

  \songline{grant a day of peace and hope}

  \songline{in Your mercy, Lord of all,}

  \songline{and remission of our sins;}

  \songline{lead us on the path of good.}

  \songlinein{May Your children never lose}

  \songline{the strong hope we place in You.}

  \songline{Lord of all, close not the door}

  \songline{in Your weak, frail children's face.}

  \songline{You know, Father, our follies}

  \songline{and the weakness of our race.}

  \songline{reward us with Your mercy;}

  \songline{look not on our little worth.}

  \songlinein{Sow on earth Your unity,}

  \songline{peace and charity, o Lord.}

  \songline{may all those who tend your sheep}

  \songline{see much good fruit in that field.}

  \songline{heal the sick, lift those who fall,}

  \songline{bestow health to one and all;}

  \songline{cleanse the sins from everyone;}

  \songline{increase holiness in all.}

  \songlinein{May Your hand save us, o Lord,}

  \songline{in all paths and ways we go,}

  \songline{as You once kept safe from Saul}

  \songline{David in his lowliness.}

  \songline{may each step, o Lord, we take}

  \songline{be according to Your will,}

  \songline{that we reach our goal in peace}

  \songline{under Your protecting care.}

  \songlinein{Through the Red Sea, You once led}

  \songline{Moses and Your chosen ones.}

  \songline{Daniel in the lion's den}

  \songline{was preserved, Lord, by Your grace.}

  \songline{Ananias and his friends}

  \songline{were saved from the furnace fire,}

  \songline{guard us also by Your power,}

  \songline{Lord, from all the evil ones.}

  \songlinein{Rising with the morning sun}

  \songline{we adore our Father's Name,}

  \songline{and sing glory to His Son,}

  \songline{as we thank the Spirit blest.}

  \songline{may our loving Father's grace}

  \songline{and the mercy of His Son}

  \songline{and the Spirit's compassion}

  \songline{be upon us all our days.}

  \songlinein{Lord, we place our hope in You}

  \songline{You the Healer of our lives.}

  \songline{may Your mercy's medicine}

  \songline{save us from the wounds of sin.}

  \songline{You alone can give us strength}

  \songline{that we need and keep Your word}

  \songline{help, guide and protect, o Christ,}

  \songline{those who worship You, o Lord.}

  \songlinein{Let us request forgiveness}

  \songline{from the Lord compassionate.}

  \songline{who waits at His open door}

  \songline{for the sinners who repent.}

  \songline{day by day we turn to You,}

  \songline{yet our sins are multiplied.}

  \songline{grant us mercy, Christ our Lord,}

  \songline{that we enter Your kingdom.}
\end{doublecolsong}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{* Fixed doxology}

\label{at-the-sunrise}

\liturgicalOption{For recital:}

\begin{halfparskip}
  At the sunrise of the morning we glorify You, Lord, for You are the Saviour of all creatures. Give us in Your compassion a day full of peace, and grant us forgiveness of sins.

  Cut not off our hope; shut not Your door in our face. Stay not Your care for us; and recompense us not as we deserve, God, for You alone know our weakness.

  Sow, our Lord, in the world love, peace and concord. Raise up priests, kings and judges, and pacify the leaders (\translationoption{nobles}). Heal the sick, preserve the healthy, and pardon the debts of all the sons of men.

  On the way on which we walk, let Your goodness keep us, my Lord, as (it kept) the boy David from Saul, give us, in Your compassion, that wherever we go, we may arrive according to Your will in peace.

  By the goodness which preserved the prophet Moses in the sea, and raised up Daniel from the den, and by which were rescued Ananias and his company in the furnace, deliver us, our Lord, from the evil one.
\end{halfparskip}

\begin{halfparskip}
  \liturgicalhint{Another:} At morning time we arise and we all worship the Father, lift up praise to the Son, and thank the Holy Spirit.

  May the loving-kindness of the Father, the compassion of the Son, and the pity of the Spirit, the threefold Mystery, help us always all our days.

  In You, our Helper, true Physician, we trust. Apply the medicine of Your mercy, heal our bruises, that we be not utterly destroyed.

  Without Your help we are very weak in keeping Your commandments. Christ, who aid those who fulfil Your will, keep Your worshippers.

  Let us beg with a fervent prayer and beseech mercy, and ask for forgiveness from the compassionate One, whose door is open to all who return to Him.

  Day by day I promise You that tomorrow I will repent. My days have passed and gone, and my trespasses remain. Christ, pity me and have mercy upon me.

  At morning time let us praise Your loving-kindness, for You have ordered the light for the joy of creation.
\end{halfparskip}

\begin{halfparskip}
  \liturgicalhint{Prayers of the Anthems of the martyrs:} \rr~Let us pray, peace be with us.

  \cc~Of You, great, glorious and true Hope, for whom the just and the prophets waited in their generations, whom the just pleased by their labours, and the holy martyrs propitiated by the precious blood which poured forth from their necks, our Lord and our God, we beseech and ask that we may find refuge in the power hidden in their bones, and be helped by their prayers, and with them, among them, in their companies, around their resting-place, and in the glorious temples set apart for the honour of their service, be made worthy to lift up to You praise, honour, confession and worship at all times, Lord of all...

  \cc~Our Lord and our God, may the martyrs, who by Your power and help overcame, humiliated and brought to nought the error of paganism, who exalted and raised the true faith of Your Name, beseech and intercede for us of Your justice, my Lord, in the great and glorious day when Your justice is revealed from heaven, just Judge, full of mercy, who forgive offences and sins, Lord of all...
\end{halfparskip}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\liturgicalhint{Prayers for help (D'udrana) when several priests are present, in order of precedence, each one is concluded by ``Amen, bless, my Lord''. The first one is always said\footnote{Assyrian da-qdam WA d'batar, p. 36: ``Only if the office takes place in a church''.}:}

\cc~1. Let, my Lord, the help of Your mercy, the support of Your grace, the hidden and glorious strength of Your glorious Trinity, and Your right hand full of mercy and compassion, overshadow and accompany the weakness of Your worshippers, from Your holy house which is full of all helps and all blessings, by the prayer of the blessed mart Mary and by the prayer of all the saints who please (\translationoption{propitiate}) You, Lord of all...

2. Our Lord and God, may Your servants be blessed with Your blessing, and Your adorers be guarded by the care of Your will; Lord, may the constant peace of Your Divinity and the enduring peace of Your Majesty reign over Your people and in Your Church, all the days of the world, Lord of all, Father...

3. May the blessing of the One who blesses all, the peace of the One who pacifies all, the mercy of the One who shows mercy on all, the protection of our adorable God, be with us, among and around us, and protect us from the evil one and from his powers at all times and hours, Lord of all, Father...

4. Our Lord and our God, may we be blessed with Your blessing, may we be protected by Your providence, may Your strength come to our support, may Your help accompany us, may Your right hand overshadow us, may Your peace reign among us, may Your Cross be a high fortress (\translationoption{wall}) and a refuge to us, and may we be defended under its wings from the evil one and his powers, at all times and hours, Lord of all, Father...

5. Lord, blessed is the mercy of Your goodness, adorable are the promises of Your Lordship, which instruct us always to look at You and to glory in You. Let not our hope be severed from You all days of our life, Lord of all, Father...

6. Our Lord and God, may Your blessing rest upon Your people, and may Your mercy continually be upon us, weak sinners; our good Hope, our merciful Refuge and Forgiver of debts and sins, Lord of all, Father...

7. May the peace of the Father be with us, and the love of the Son among us, and may the Holy Spirit lead us according to His will, and let His mercy and compassion be on us at all times and hours, Lord of all, Father...

8. Lord, may Your peace dwell in us, and Your tranquillity rule in us, and may Your charity increase among us all the days of our lives, Lord of all, Father...

9. Lord, protect us by Your right hand, defend us under Your wings, and let Your help accompany us all the days of our lives, Lord of all, Father...

10. Lord, give to us Your constant peace, charity, love of learning, life, happiness and joy; and do not let us lack Your sustenance on any day of our lives, Lord of all, Father...

11. Be a sleepless guard to the stronghold in which dwell Your sheep, that they may not be hurt by the wolves that are thirsting for the blood of Your flock: for You are the sea which will not decrease, Lord of all, Father...

12. Lord, bless us with Your blessings, surround us with the fortress (\translationoption{wall}) of Your care, do not deprive us of good things and make us recline at table in Your luminous bridal feast, Lord of all, Father...

13. Lord, come to our help in Your mercy, reveal our redemption in Your compassion, and direct our steps through the paths of justice all the days of our lives, Lord of all, Father...

14. Lord, let Your goodness dawn on us when Your justice judges us and let Your mercy come to our help on the day on which Your Majesty will appear, Lord of all, Father...

15. Lord of all, make Your blessing and Your grace, Your right hand full of mercy and compassion, overshadow and accompany the communities of Your adorers who call and beseech You at all times and hours, Lord of all, Father...

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Final prayers}

\liturgicalhint{Long version:}

\liturgicalhint{Of Mary.} May the prayer, O my Lord, of the holy virgin, the petition of the blessed mother, the beseeching and entreating of her who is full of grace, mart Mary the blessed, the great power of the victorious Cross, the divine help, and the supplication of mar John the Baptist be with us continually, at all seasons and times, Lord of all, etc.~--- \rr~Amen. Bless, o my Lord.

\liturgicalhint{Of the Apostles.} May the prayer, O my Lord, of the holy apostles, the petition of the true preachers, the beseeching and entreating of the triumphant (\translationoption{illustrious}) athletes, the preachers of righteousness, the announcers of sanctity and sowers of peace in creation, be with us continually, at all seasons and times, Lord of all, etc.~--- \rr~Amen. Bless, o my Lord.

\liturgicalhint{Of the saints.} May the prayer, petition, beseeching and entreating of our noble and holy father mar Thomas the blessed apostle, of mar Adai and mar Mari, teachers of the East, of mar Stephen the firstborn of the martyrs, of mar Simeon bar Sabae, of mar Jacob, of mar Ephrem, of the giant of strength, mar George the illustrious martyr, of mar Quriaqos, of mar Pethion and mar Hormizd, of blessed mar Augin and of all his spiritual company, of saint Barbara and Shmuni and her sons, of Meskenta and her two sons, and of all martyrs and saints of our Lord, be with us continually, a high wall and strong house of refuge, to save, deliver, rescue, and guard our bodies and souls from the evil one and his hosts, at all seasons and times, Lord of all...~--- \rr~Amen. Bless, o my Lord.

\liturgicalhint{Short version:}

\cc~May the prayer, my Lord, of the holy virgin, the petition of the blessed mother, the beseeching and entreating of her who is full of grace, mart Mary, the blessed; the great power of the victorious Cross, the divine help, the supplication of mar John Baptist, the prayer of the holy apostles, the petition of the true preachers, the beseeching and entreating of the triumphant athletes, the preachers of righteousness, the announcers of sanctity and sowers of peace in creation; the intercession of our father mar Thomas the blessed apostle, of mar Adai and mar Mari, teachers of the East, of mar Stephen, the firstborn of the martyrs, of mar Simeon bar Sabae, of mar Jacob, of mar Ephrem, of the giant of strength, mar George the illustrious martyr, of mar Quriaqos, of mar Pethion and mar Hormizd, of blessed mar Augin and of all his spiritual company, of saint Barbara and Shmuni and her sons, of Meskenta and her two sons, and of all martyrs and saints of our Lord, be with us continually, a high wall and strong house of refuge, to save, deliver, rescue, and guard our bodies and souls from the evil one and his hosts, at all seasons and times, Lord of all, etc.~---

[\liturgicalhint{Assyrian: The priest takes the cross in his hand, and turning to the people says:} Bless, my Lord; by your command. \liturgicalhint{They answer,} By the command of Christ and glory to his holy Name. \liturgicalhint{And they bow their heads.}]

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\markedsection{Hutama}

Bestow, our Lord and our God, in Your loving-kindness, at this morning time, salvation to the oppressed, release to the prisoners, restoration to the tormented, healing to the sick, return to those who are far away, preservation to those who are near, pardon to sinners, acceptance to the penitent, exaltation to the righteous, supply of their needs to the poor, finding to the lost, bringing back to them that are driven away, a good and acceptable memorial to the departed, mercy and compassion to all that are created and made; and grant to us and all men those things which will help us, weak sinners, and please Your Majesty, in Your loving-kindness and mercy, Christ, the hope of our nature, now and at all times and for ever and ever.

\liturgicalhint{Other Hutame:}

1. Glory to You, Jesus our victorious King, the Radiance of the eternal Father, begotten without beginning, above all times and (created) beings; we have no hope and expectation unless it be You, the Creator. By the prayer of the just and elect who pleased You from the beginning, pardon our sins, forgive our offences, deliver us from our afflictions, answer our requests, bring us to Your glorious light, and guard us by Your living Cross from all harm, hidden and open, now and at all times and for ever and ever.

2. May the prayer of Your feeble servants, our Lord and our God, be received before the throne of Your Godhead; and may this our assembling please the will of Your Majesty; that we may receive from You the gift of health of the body and safe keeping of the soul, superabundance of sustenance (\translationoption{food}), remission of debts, pardon of sins, constant peace, O my Lord, and prolonged prosperity; unity of love which passes not away and departs not from among us, in every age of this world, now and at all times and for ever and ever.

3. Blessed is God forever and glorified is His holy Name in all eternity. To Him we make request, and of the overflowing sea of His mercy we beseech, that He will make us worthy of the excellent glory of His kingdom, of felicity with His holy angels, of confidence (\translationliteral{openness of face}) before Him, and of standing at His right hand in Jerusalem which is above, in His grace and mercy, now and at all times and forever and ever.

4. May God the Lord of all, in whose house we have assembled, and before whose Majesty we have prayed, in the hope of His great mercy, hear our prayer in His compassion, receive our request in His pity; may He wash and cleanse the filth of our trespasses and sins with the hyssop of His abundant pitifulness, may He give rest to the souls of our departed in the glorious mansions of His kingdom; may He sprinkle us all with the dew of His sweetness, and may the right hand of His care overshadow us and all creatures in His grace and mercy, now and at all times and forever and ever.

5. God the Lord of all, who has entrusted His praises to our mouth, His songs to our tongue, His canticles to our throats, His confession to our lips, His faith to our hearts, may He hear our prayers, accept our requests, be pleased with our beseechings, ease our debts, answer our petitions with benefits and not with blame; and from the great treasury of His mercy may He pour forth His mercy and compassion on us and on the whole world, now and at all times and forever and ever.

6. May the Name of God the Lord of all, who orders times and seasons, be glorified among us; may the right hand of the care of His mercy overshadow us who are feeble and sinful, and the whole world, the holy Church and its children; our fathers, brothers, superiors and teachers; our departed who were separated from us and removed from among us, and all our brotherhood in Christ, now and at all times and forever and ever.

7. To God be glory; to the angels honour; to Satan shame; to the Cross worship; to the Church exaltation; to the departed resurrection; to the penitent acceptance; to prisoners release; to the sick and infirm recovery and healing; to the four quarters of the world great peace and prosperity. Also on us, who are weak and sinful, may the compassion and mercy of our adorable God come, may they overshadow us, be poured forth, remain and reign constantly, now and at all times and for ever and ever.

\liturgicalhint{Assyrian Hudra:}

8. By the right hand of Your majesty, our Father in heaven, bless us all, my Lord, keep us all, help us all, assist and protect us all, quicken all our departed, let Your right hand overshadow us all, may Your mercy and compassion be poured forth on us all, and may continual praise, honour, confession, worship, thanksgiving ascend to You from the mouth of us all, Christ, hope of our nature, now and at all times and forever and ever.

9. \liturgicalhint{In a monastery or house:} May God the Lord of all bless our congregation in His grace, and keep us from falling in the overflowing multitude of His mercy; may He answer our petitions from His treasury; and on all the whole world, on the holy Church and its children, on this country and its inhabitants, on this habitation and those who dwell in it, and on us all and everyone of us together, may the compassion and mercy of our good God come and be poured forth continually, Christ the hope of our nature, now and at all times and forever and ever.

[\liturgicalhint{Hudra:} \rr~May Christ hear your prayers; may Christ make your priesthood famous in the heavenly kingdom.]

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\end{document}